---
title: "SF Bikeshare Status - Modeling Setup"
output: html_notebook
---


**NOTE:** before running this chunk, the AWS Instance Type was switched from t2.micro to t3.2xlarge.

**NOTE:** This should be run before each new model to create the relevant dataset for modeling:  lmtd_vars_lag3.



## Setup    
  Load the relevant libraries.
```{r}
# {r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


library("tidyverse")
library("data.table")
library("caret")

```
  
    
  Session Info.
```{r}

sessionInfo()

```


  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/SF_Bikeshare_Status/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
  
    
  Get the base dataset.
```{r}

status_trips_time <-
  read_rds(path = paste0(wd,
                         "/Data/Interim/",
                         "status_trips_time.rds"
                         )
           )

message("status_trips_time")
str(status_trips_time)
# View(status_trips_time %>% head(1000))

```
  
    
  Split the dataset by `station_id` and cleanup the resulting dataset.
```{r}

# initial split
id_split <-
  status_trips_time %>% 
  split(.$station_id)


# data cleanup
id_split <-
  pmap(.l = list(a = id_split),
       .f = function(a) {
         df = a %>% 
           tail(n = -3) %>% # remove first three rows because of lags
           mutate_if(is.factor, factor) %>% 
           select(-docks_avail_now) %>% # remove `docks_avail_now` because you won't know this at the time of prediction
           as.data.table() %>% 
           setkey(time_rnd, station_id)
         
         return(df)
         }
       )

```
  
    
  Create splits for train/valid/test.
```{r}

# create a function for the split
func_train_valid_test <-
  function(data, pct_train, pct_valid, ...) {
    row_cnt = data %>% nrow()
    
    pct_train = pct_train
    pct_valid = pct_valid
    pct_test = 1 - pct_train - pct_valid
    
    split_pt_train = round(pct_train * row_cnt, digits = 0)
    split_pt_valid = round((pct_train + pct_valid) * row_cnt, digits = 0)
    
    dt =
      data %>% 
      arrange(time_rnd) %>% 
      mutate(row_num = row_number(),
             data_type = case_when(row_num < split_pt_train ~ "01_train",
                                   between(x = row_num,
                                           lower = split_pt_train,
                                           upper = split_pt_valid,
                                           incbounds = TRUE
                                           ) ~ "02_valid",
                                   TRUE ~ "03_test"
                                   ) %>% 
               factor()
             )
    }


# run the function
id_split <-
  pmap(.l = list(a = id_split),
       .f = function(a) {
         data = func_train_valid_test(data = a, pct_train = 0.70, pct_valid = 0.15)
         }
       )


# confirmation
pmap(.l = list(a = id_split,
               b = names(id_split)
               ),
     .f = function(a, b) {
       message(b)
       
       str(a)
       }
     )

# View(id_split$`10` %>% head(1000))

```
  
    
  Save the data.
```{r}

# write_rds(id_split,
#           path = paste0(wd,
#                         "/Data/Interim/",
#                         "id_split.rds"
#                         )
#           )

```
  
    
  Limit the varaibles to a smaller amount to speed up calculations.
```{r}

small_test <-
  id_split %>%
  map(~select(.x,
              row_num,
              data_type,
              time_rnd,
              station_id,
              bikes_avail_now,
              bikes_avail_lag05,
              bikes_avail_lag10,
              bikes_avail_lag15,
              wday.lbl,
              hour,
              minute
              )
      )

glimpse(small_test$`10`)

```
  
    
  One-hot enconde the data.
```{r}

one_hot <-
  pmap(.l = list(a = small_test),
       .f = function(a) {
         df = select(a,
                     -station_id,
                     -data_type,
                     -time_rnd,
                     -row_num
                     )
         
         dmy = dummyVars(bikes_avail_now ~ ., data = df)
         
         return(dmy)
         }
       )


id_split_one_hot <-
  pmap(.l = list(a = one_hot,
                 b = id_split
                 ),
       .f = function(a, b) {
         b %>%
           select(station_id,
                  data_type,
                  bikes_avail_now,
                  time_rnd,
                  row_num
                  ) %>% 
           bind_cols(predict(object = a,
                             newdata = b %>% 
                               select(-station_id,
                                      -data_type,
                                      -time_rnd,
                                      -row_num
                                      )
                             ) %>% 
                       as.data.frame()
                     ) %>% 
           arrange(time_rnd)
         }
       )

class(id_split_one_hot$`10`)
glimpse(id_split_one_hot$`10`)
# View(id_split_one_hot$`10` %>% head(100))


rm(one_hot)

```
  
    
  Create lags of 3.
```{r}

vars_to_lag_dt <-
  id_split_one_hot %>% 
  map(~select(.x, wday.lbl.Friday:minute.55) %>%
        as.data.table()
      )

vars_to_not_lag_dt <-
  id_split_one_hot %>% 
  map(~select(.x,
              station_id:bikes_avail_lag15,
              ) %>% 
        as.data.table()
      )

# id_split %>% map(~colnames(.x))
# vars_to_lag_dt %>% map(~colnames(.x))

lags_3 <-
  pmap(.l = list(a = vars_to_lag_dt),
       .f = function(a) {
         dat = a[ , sapply(names(a),
                           function(x) {
                             paste0(x, '_lag', c("05min", "10min", "15min"))
                             }
                           ) := shift(.SD, 1:3),
                  .SDcols = wday.lbl.Friday:minute.55
                  ][]
         
         dat = dat[4:nrow(a), ]
         
         return(dat)
         }
       )
# lags_3$`10` %>% select(matches("minute\\.0")) %>% head(100) %>% View()

lmtd_vars_lag3 <-
  pmap(.l = list(a = vars_to_not_lag_dt,
                 b = lags_3
                 ),
       .f = function(a, b) {
         bind_cols(a[4:nrow(a), ],
                   b
                   ) %>% 
           select(-(wday.lbl.Friday:minute.55)
                  )
         }
       )

# lmtd_vars_lag3 %>% map(~colnames(.x))
lmtd_vars_lag3$`10` %>% glimpse()
# View(lmtd_vars_lag3$`10` %>% 
#        select(time_rnd,
#               matches("month\\.lbl\\.August")
#               ) %>% 
#        head(100)
#      )
# View(lmtd_vars_lag3$`10` %>% 
#        select(time_rnd,
#               matches("minute\\.0")
#               ) %>% 
#        head(100)
#      )
((lmtd_vars_lag3$`10` %>% ncol()) - 5) / 3 # remove 5 non-lagged variables, and dividing by 3 gives the total number of vars without ags


rm(vars_to_lag_dt, vars_to_not_lag_dt, lags_3)

```
  
    
  Save the data.
```{r}

# write_rds(lmtd_vars_lag3,
#           path = paste0(wd,
#                         "/Data/Interim/",
#                         "lmtd_vars_lag3.rds"
#                         )
#           )

```  
  
    
  Remove no-longer-needed files.
```{r}

rm(id_split, id_split_one_hot, small_test, status_trips_time, func_train_valid_test)

```
  

**NOTE:** Don't forget to switch the AWS Instance Type was switched from t3.2xlarge. back to t2.micro.


