---
title: "SF Bikeshare Status - Initial Explore"
output: html_notebook
---


  
## Setup    
  Load the relevant libraries.
```{r}
# {r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


library("tidyverse")
# install.packages("data.table")
library("data.table")
library("lubridate")
library("DBI")
# install.packages("RSQLite")
library("RSQLite")

```
  
    
  Session Info.
```{r}

sessionInfo()

```


  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/SF_Bikeshare_Status/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
  
    
  Get the data from the SqlLite database.
```{r}

# connect to the sqlite database
con <- dbConnect(SQLite(),
                 paste0(wd,
                        "/Data/Raw/",
                        "database.sqlite"
                        )
                 )


# get a list of the talbles
as.data.frame(dbListTables(con))


# get data for every 5 minutes
myQuery <- dbSendQuery(con,
                       "Select          *

                        From            status

                          Where         (     time like '%:_0:__'
                                          or  time like '%:_5:__'
                                         )
                       "
                       )

sf_status <- dbFetch(myQuery, n = -1) %>% 
  as.data.table


class(sf_status)
glimpse(sf_status)
object.size(sf_status)
summary(sf_status)

dbClearResult(myQuery)
# View(sf_status %>% head(500))

```
  
    
  For testing purposes, select only six `station_id` values - the two most frequently used stations, the two least frequently used stations, and two in the middle.
```{r}

avg_bikes_available <-
  sf_status[ ,
            .(bike_avail_cnt = .N,
              bike_avail_avg = mean(bikes_available, na.rm = TRUE)
              ),
            by = station_id
            ][order(bike_avail_cnt,
                    bike_avail_avg
                    )
              ][ , row_num := row_number(bike_avail_cnt)]

# View(avg_bikes_available)


station_id_2top <-
  avg_bikes_available %>% 
  top_n(n = 2,
        wt = row_num
        ) %>% 
  pull(station_id)

station_id_2bot <-
  avg_bikes_available %>% 
  top_n(n = -2,
        wt = row_num
        ) %>% 
  pull(station_id)

station_id_2mid <-
  avg_bikes_available %>% 
  filter(row_num == nrow(avg_bikes_available) / 2 |
           row_num == (nrow(avg_bikes_available) / 2) + 1
         ) %>% 
  pull(station_id)


station_id_select <-
  c(station_id_2top, station_id_2bot, station_id_2mid)

station_id_select


rm(station_id_2top, station_id_2bot, station_id_2mid)

```
  
    
  Save the data.
```{r}

write_rds(station_id_select,
          path = paste0(wd,
                        "/Data/Interim/",
                        "station_id_select.rds"
                        )
          )

```
    
    
  Update variable types & confirm data pull.
```{r}

sf_status_select_ids <-
  sf_status[station_id %in% station_id_select]

sf_status_select_ids$time <- as_datetime(sf_status_select_ids$time)
sf_status_select_ids$time_min <- minute(sf_status_select_ids$time)

setkey(sf_status_select_ids, time_min)

class(sf_status_select_ids)
glimpse(sf_status_select_ids)
object.size(sf_status_select_ids)
summary(sf_status_select_ids)

unique(sf_status_select_ids$time_min)

sf_status_select_ids <-
  sf_status_select_ids[ ,time_min := NULL]

str(sf_status_select_ids)

```
  
    
  Save the data.
```{r}

write_rds(sf_status_select_ids,
          path = paste0(wd,
                        "/Data/Interim/",
                        "sf_status_select_ids.rds"
                        )
          )

```
  
    
  Remove no-longer-needed files.
```{r}

rm(avg_bikes_available, con, myQuery, sf_status)

```
  

