---
title: "SF Bikeshare Status - Initial Explore"
output: html_notebook
---


  
## Setup    
  Load the relevant libraries.
```{r}
# {r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


library("tidyverse")
# install.packages("data.table")
library("data.table")
library("lubridate")
library("DBI")
# install.packages("RSQLite")
library("RSQLite")

```
  
    
  Session Info.
```{r}

sessionInfo()

```


  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/SF_Bikeshare_Status/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
  
    
  Get the `station_id` values produced in `obtain_base_data.Rmd`.
```{r}

station_id_select <-
  read_rds(path = paste0(wd,
                         "/Data/Interim/",
                         "station_id_select.rds"
                         )
           )

station_id_select

```
  
    
  Get the data from the SqlLite database.
```{r}

# connect to the sqlite database
con <- dbConnect(SQLite(),
                 paste0(wd,
                        "/Data/Raw/",
                        "database.sqlite"
                        )
                 )


# get a list of the talbles
as.data.frame(dbListTables(con))


# get data for every 5 minutes
myQuery <- dbSendQuery(con,
                       "Select          *

                        From            trip

                        Where         start_station_id in (22, 61, 84, 83, 10, 21)
                            Or        end_station_id in (22, 61, 84, 83, 10, 21)
                       "
                       )

sf_trip <- dbFetch(myQuery, n = 10000) %>% 
  as.data.table


class(sf_trip)
glimpse(sf_trip)
object.size(sf_trip)
summary(sf_trip)

dbClearResult(myQuery)
# View(sf_trip %>% head(500))

rm(con, myQuery)

```
  
    
  Update variable types.
```{r}

sf_trip_var_type_update <-
  sf_trip %>% 
  mutate_at(c("id", "start_station_name", "start_station_id", "end_station_name",
              "end_station_id", "bike_id", "subscription_type", "zip_code"
              ),
            as.factor
            ) %>% 
  mutate(start_date = paste0(start_date, ":00"),
         end_date = paste0(end_date, ":00")
         )


func_chr_to_dt <-
  function(data, col) {
    col_enquo = enquo(col)
    
    new_df =
      data %>% 
      mutate(mth = str_extract(string = !!col_enquo, pattern = "^\\d{1,2}") %>% 
               str_pad(width = 2, side = "left", pad = "0"),
             day = str_extract(string = !!col_enquo, pattern = "(?<=/)\\d{1,2}(?=/)") %>% 
               str_pad(width = 2, side = "left", pad = "0"),
             yr = str_extract(string = !!col_enquo, pattern = "(?<=/)\\d{4}(?=\\s)"),
             hr = str_extract(string = !!col_enquo, pattern = "(?<=\\s)\\d{1,2}(?=:)") %>% 
               str_pad(width = 2, side = "left", pad = "0"),
             min = str_extract(string = !!col_enquo, pattern = "(?<=:)\\d{1,2}(?=:)") %>% 
               str_pad(width = 2, side = "left", pad = "0"),
             sec = str_extract(string = !!col_enquo, pattern = "(?<=:)\\d{1,2}$") %>% 
               str_pad(width = 2, side = "left", pad = "0"),
             dt_chr = paste0(yr, "-", mth, "-", day, " ", hr, ":", min, ":", sec),
             dt = as_datetime(dt_chr, tz = "America/Los_Angeles")
             ) %>% 
      select(-(mth:dt_chr))
    
    return(new_df)
    }


sf_trip_clean_dates <-
  func_chr_to_dt(data = sf_trip_var_type_update,
                 col = start_date
                 ) %>% 
  rename(start_dt = dt) %>% 
  func_chr_to_dt(col = end_date) %>% 
  rename(end_dt = dt)

str(sf_trip_clean_dates)
summary(sf_trip_clean_dates)
# View(sf_trip_clean_dates)


rm(sf_trip, sf_trip_var_type_update)

```
  
    
  Create `date_ceiling` variables to aggregate at the 5min level.
```{r}

date_ceilings <-
  sf_trip_clean_dates %>% 
  mutate(start_dt_5min_ceil = ceiling_date(start_dt, unit = "5 mins"),
         end_dt_5min_ceil = ceiling_date(end_dt, unit = "5 mins")
         )

str(date_ceilings)
# View(date_ceilings)


rm(sf_trip_clean_dates)

```
  
    
  Create a "wide" counts dataset to later be joined to the station status data.
```{r}

# Initial counts
cnts_start <-
  date_ceilings %>% 
  filter(start_station_id %in% station_id_select) %>% 
  group_by(start_station_id, subscription_type, start_dt_5min_ceil) %>% 
  summarise(cnt_5min_window_start = n()
            ) %>% 
  ungroup() %>% 
  arrange(start_station_id, subscription_type, start_dt_5min_ceil)


cnts_end <-
  date_ceilings %>% 
  filter(end_station_id %in% station_id_select) %>% 
  group_by(end_station_id, subscription_type, end_dt_5min_ceil) %>% 
  summarise(cnt_5min_window_end = n()
            ) %>% 
  ungroup() %>% 
  arrange(end_station_id, subscription_type, end_dt_5min_ceil)

# View(cnts_start)
# View(cnts_end)

cnts_start_end <-
  cnts_start %>% 
  full_join(y = cnts_end,
            by = c("start_station_id" ="end_station_id",
                   "subscription_type" = "subscription_type",
                   "start_dt_5min_ceil" = "end_dt_5min_ceil"
                   )
            ) %>% 
  arrange(start_station_id, start_dt_5min_ceil) %>% 
  rename(station_id = start_station_id,
         dt_5min_ceil = start_dt_5min_ceil
         ) %>% 
  mutate(station_id = factor(station_id))

str(cnts_start_end)
summary(cnts_start_end)
# View(cnts_start_end %>% head(1000))


# turn into a wide dataset
trip_cnts_start_end_wide <-
  cnts_start_end %>% 
  gather(key = "key",
         value = "value",
         starts_with("cnt_5min_window")
         ) %>% 
  unite(col = new,
        subscription_type,
        key
        ) %>% 
  spread(key = new,
         value = value,
         fill = 0
         ) %>% 
  # ungroup() %>% 
  as.data.table() %>% 
  setkey(station_id, dt_5min_ceil)

str(trip_cnts_start_end_wide)
View(trip_cnts_start_end_wide)



# min(cnts_start_end_wide$dt_5min_ceil)
# max(cnts_start_end_wide$dt_5min_ceil)
# 
# 
# 
# all_dt <-
#   data.frame(dt = seq.POSIXt(from = as_datetime("2013-08-29 00:00:00",
#                                                 tz = "America/Los_Angeles"
#                                                 ),
#                              to = as_datetime("2014-01-23 00:00:00",
#                                               tz = "America/Los_Angeles"
#                                               ),
#                              by = "5 mins"
#                              )
#              ) %>% 
#   as.data.table() %>% 
#   setkey(dt)
# 
# str(all_dt)
# View(all_dt %>% head(1000))
# 
# 
# all_dt_all_cnts <-
#   all_dt %>% 
#   left_join(y = cnts_start_end_wide,
#             by = c("dt" = "dt_5min_ceil")
#             ) %>% 
#   filter(!is.na(station_id)) %>% 
#   as.data.table() %>% 
#   setkey(station_id, dt)
# 
# str(all_dt_all_cnts)
# View(all_dt_all_cnts %>% head(1000))

```
  
    
  Save the data.
```{r}

write_rds(trip_cnts_start_end_wide,
          path = paste0(wd,
                        "/Data/Interim/",
                        "trip_cnts_start_end_wide.rds"
                        )
          )

```
  
    
  Remove no-longer-needed files.
```{r}

rm(cnts_end, cnts_start, cnts_start_end, date_ceilings, func_chr_to_dt)

```
  

