---
title: "SF Bikeshare Status - Create Single Dataset"
output: html_notebook
---


  
## Setup    
  Load the relevant libraries.
```{r}
# {r, message=FALSE, warning=FALSE}

# rm(list = ls())
# .rs.restartR()


library("tidyverse")
# install.packages("data.table")
library("data.table")
library("lubridate")
# library("DBI")
# install.packages("RSQLite")
# library("RSQLite")

```
  
    
  Session Info.
```{r}

sessionInfo()

```


  Setup the root directory.
```{r "setup", include = FALSE}

require("knitr")

opts_knit$set(root.dir = "/home/rstudio/Dropbox/_AWS/SF_Bikeshare_Status/")

```
  
    
  Setting `wd` as the working directory.
```{r}

wd <- getwd()

wd

```
  
    
  Get the base datasets..
```{r}

station_id_select <-
  read_rds(path = paste0(wd,
                         "/Data/Interim/",
                         "station_id_select.rds"
                         )
           )

message("station_id_select")
str(station_id_select)


sf_status_select_ids <-
  read_rds(path = paste0(wd,
                         "/Data/Interim/",
                         "sf_status_select_ids.rds"
                         )
           )

message("sf_status_select_ids")
str(sf_status_select_ids)
# View(sf_status_select_ids %>% arrange(station_id, time))


trip_cnts_start_end_wide <-
  read_rds(path = paste0(wd,
                         "/Data/Interim/",
                         "trip_cnts_start_end_wide.rds"
                         )
           )

message("trip_cnts_start_end_wide")
str(trip_cnts_start_end_wide)
# View(trip_cnts_start_end_wide %>% arrange(station_id, dt_5min_ceil))

```
  
    
  To prepare the datasets for joinig, round `sf_status_select_ids$time`, and update the variable types for the variables on which the data are joined.
```{r}

# round to nearest minute
sf_status_select_ids <-
  sf_status_select_ids[ , time_rnd := floor_date(time, unit = "minutes")][ , time := NULL]


# update variable types
sf_status_select_ids$station_id <- as.character(sf_status_select_ids$station_id)
trip_cnts_start_end_wide$station_id <- as.character(trip_cnts_start_end_wide$station_id)

sf_status_select_ids$time_rnd <- as.character(sf_status_select_ids$time_rnd)
trip_cnts_start_end_wide$dt_5min_ceil <- as.character(trip_cnts_start_end_wide$dt_5min_ceil)


message("sf_status_select_ids")
str(sf_status_select_ids)
# View(sf_status_select_ids %>% arrange(station_id, time_rnd) %>% head(1000))


message("trip_cnts_start_end_wide")
str(trip_cnts_start_end_wide)
# View(trip_cnts_start_end_wide %>% arrange(station_id, dt_5min_ceil) %>% head(1000))

```
  
    
  Merge the datasets, replace `NA` with zero, and update variable types.
```{r}

# merge itself
status_trips <-
  merge(x = sf_status_select_ids,
        y = trip_cnts_start_end_wide,
        by.x = c("station_id", "time_rnd"),
        by.y = c("station_id", "dt_5min_ceil"),
        all.x = TRUE
        )

# str(status_trips)
# View(status_trips %>% head(1000))
# summary(status_trips)


# NA to 0
status_trips[is.na(status_trips)] <- 0


# update variable type
status_trips <-
  status_trips %>% 
  mutate(station_id = as.factor(station_id),
         time_rnd = as_datetime(status_trips$time_rnd, tz = "America/Los_Angeles")
         ) %>% 
  mutate_at(vars(matches("cnt_5min")), as.integer) %>% 
  as.data.table() %>% 
  setkey(station_id, time_rnd)


str(status_trips)
View(status_trips %>% head(1000))
summary(status_trips)

```
  
    
  Save the data.
```{r}

write_rds(status_trips,
          path = paste0(wd,
                        "/Data/Interim/",
                        "status_trips.rds"
                        )
          )

```
  
    
  Remove no-longer-needed files.
```{r}

rm(sf_status_select_ids, trip_cnts_start_end_wide, station_id_select)

```
  

